---
title: "Regression Analytics"
author: "Dev"
date: "2023-11-04"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,comment = NA)
```

## R Markdown

**Q-1:Run the following code in R-studio to create two variables X and Y.**
**set.seed(2017)**
**X=runif(100)*10**
**Y=X*4+3.45**
**Y=rnorm(100)*0.29*Y+Y**

```{r}
set.seed(2017)
X=runif(100)*10
Y=X*4+3.45
Y=rnorm(100)*0.29*Y+Y
```

**a) Plot Y against X. Include a screenshot of the plot in your submission. Using the file menu you can save the graph as a picture on your computer. Based on the plot do you think we can fit a linear model to explain Y based on X? (8% of total points)**

```{r}
plot(X,Y,xlab="Frequency of X",ylab = "Frequency of Y", main = "Plot of Y against X",col="red",abline(lm(Y~X),col="blue"))
```

We can clearly see that the above graph shows a positive linear relationship between X and Y.Most of the data points follow the linear model line. However, there are some outliers but majority of the points follow a positive linear regression. Hence, we can conclude that we can in fact use a linear model to explain relationship of X and Y. 

**b) Construct a simple linear model of Y based on X. Write the equation that explains Y based on X. What is the accuracy of this model? (8% of total points)**

```{r}
model=lm(Y~X)
```
We can now get coefficient and intercept of relation between X and Y as:
```{r}
round(model$coefficients,2)
```
Hence, we can write equation of Y and X as follows:
$$Y=b_nX+b_o$$
$$=>Y=3.61X+4.47$$

Accuracy of above model is :
```{r}
summary(model)
```
The accuracy of model is the value of multiple R-squared. Hence, our model is 65% accurate. A smaller p value here shows us that the relationship is significant between both variables. This is also evident from smaller difference between multiple- squared ad adjusted r-squared.

**c) How the Coefficient of Determination, R 2, of the model above is related to the correlation coefficient of X and Y? (8% of total points)**

In a linear regression model square of correlation coefficient and Multiple R-squared are equal.

Value of Correlation coefficient can be calculated as: 
```{r}
cor(X,Y)^2
```

Hence, value of correlation coefficient and R-square values are same as **0.65**.

***

**Q-2:We will use the ‘mtcars’ dataset for this question. The dataset is already included in your R distribution. The dataset shows some of the characteristics of different cars.**
```{r}
#Printing head of dataset to have a glance at it:
head(mtcars)
```

**a) James wants to buy a car. He and his friend, Chris, have different opinions about the Horse Power (hp) of cars. James think the weight of a car (wt) can be used to estimate the Horse Power of the car while Chris thinks the fuel consumption expressed in Mile Per Gallon (mpg),is a better estimator of the (hp). Who do you think is right? Construct simple linear models using mtcars data to answer the question. (17% of total points).**

Making James model-
```{r}
james=lm(mtcars$hp~mtcars$wt,data = mtcars)
plot(mtcars$wt,mtcars$hp,col="red",abline(lm(mtcars$hp~mtcars$wt)))
summary(james)
```

Making Chris model-
```{r}
chris=lm(mtcars$hp~mtcars$mpg,data = mtcars)
plot(mtcars$mpg,mtcars$hp,col="red",abline(lm(mtcars$hp~mtcars$mpg)))
summary(chris)
```

Value of R-squared is larger for Chris's opinion which means he has accuracy,i.e,
**Chris's model is more accuracte and correct.**

**b) Build a model that uses the number of cylinders (cyl) and the mile per gallon (mpg) values of a car to predict the car Horse Power (hp). Using this model, what is the estimated Horse Power of a car with 4 cylinders and mpg of 22? (17% of total points)**

```{r}
model.2=lm(hp~cyl+mpg,data=mtcars)
df=data.frame(cyl=4,mpg=22)
prediction=predict(model.2,newdata = df)
prediction
```

Hence, according to created model value of horse power would be **88.93618**.

**Q-3:The dataset contains information about houses in different parts of Boston. Details of the dataset is explained here. Note the dataset is old, hence low house prices!**

Using required dataset:
```{r}
library(mlbench)
data(BostonHousing)
```

**a) Build a model to estimate the median value of owner-occupied homes (medv)based on the following variables: crime crate (crim), proportion of residential land zoned for lots over 25,000 sq.ft (zn), the local pupil-teacher ratio (ptratio) and weather the whether the tract bounds Chas River(chas). Is this an accurate model? (Hint check R 2 ) (8% of total points)**
Making the required linear model:
```{r}
model.3=lm(medv~crim+zn+ptratio+chas,BostonHousing)
summary(model.3)
```

All variables here are significant but lower value of multiple-R squared tells us that model has a very low accuracy of 35.99%. Hence, it is not an accurate model.

**b) Use the estimated coefficient to answer these questions?**

Displaying coefficients for reference:
```{r}
model.3$coefficients
```

**I. Imagine two houses that are identical in all aspects but one bounds the Chas River and the other does not. Which one is more expensive and by how much? (8% of total points)** 

Value of coefficient for chas1 is positive. This means that the house that bounds the chas river will be more expensive by 4.58392591. 

**II. Imagine two houses that are identical in all aspects but in the neighborhood of one of them the pupil-teacher ratio is 15 and in the other one is 18. Which one is more expensive and by how much? *(Golden Question: 4% extra)**

Value of coefficient for pupil-teacher ratio is negative. This means that the house with a higher pupil-teacher ratio (18) will have a lesser price than the house with a lowere pupil-teacher ratio(15) by : 
```{r}
18*1.49367255-15*1.49367255
```

**c) Which of the variables are statistically important (i.e. related to the house price)? Hint: use the p-values of the coefficients to answer. (8% of total points)**

All of the variables are statistically important since all of them have a low p-value i.e below 0.05

**d) Use the anova analysis and determine the order of importance of these four variables. (18%of total points)**
```{r}
anova(model.3)
```
Order of importance can be found on the basis of values of Sum sq. Hence, the order of important can be deduced out of anova analysis as: 

**crim**
**ptratio**
**zn**
**chas**


